---
title: "INST 314 - Week 6B Lab - Spring 2021"
author: "Alec Mattu"
output: html_document
---
QUESTION 1
Show your code for uploading and subsetting the data to create a new dataframe TO respond to question 1.

##### Download the Atheism Dataset
```{r}
download.file("http://www.openintro.org/stat/data/atheism.RData", destfile = "~/atheism.RData")
load("~/atheism.RData")
```

##### Create a dataframe subset for Spain (ES)
```{r}
ES2012 <- subset(atheism, nationality == "Spain" & year == "2012")
summary(ES2012)
```

##### Remove empty levels
```{r}
ES2012$nationality <- factor(ES2012$nationality)
table(ES2012$nationality, ES2012$response)
```

ANSWER 1A:    Because we have over 10 successes and 10 failures (Atheists vs. Non-atheists), we are in the "clear" to use the `inference` function. See below.

Draw an inference from the dataset above.
```{r}
inference(ES2012$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
```
Textual representation of the above information, where success is considered an atheist for simplicity sake:
```{r}
## Single proportion -- success: atheist 
## Summary statistics: p_hat = 0.09 ;  n = 1145 
## Check conditions: number of successes = 103 ; number of failures = 1042 
## Standard error = 0.0085 
## 95 % Confidence interval = ( 0.0734 , 0.1065 )
```

Alternatively, we could use `prop` (`test`/`table`) as demonstrated in the Week 6 demo. Included here for posterity:
```{r}
prop.test(50, n = 1145, conf.level = .95)$conf.int # n=1145, 1145 respondents
prop.table(table(ES2012$nationality, ES2012$response))
```

ANSWER 1B:   Null Hypothesis: p_hat of 0.10 (10%). Alt Hypothesis: p_at of something other than 0.10 (!= 10%). This is a one tailed test. Because a decent confidence interval (CI) overlap exists for the 2005/2012 samples respectively, and p_hat for 2012 is within the 2005 confidence interval (CI) of (0.083 - 0.1177) sitting at ~0.09, it's unlikely that this change is anything other than chance/minor variance. See below.

Create a dataframe for Spain 2005
```{r}
ES2005 <- subset(atheism, nationality == "Spain" & year == "2005") # Create a dataframe for 2005 Spain
ES2005$nationality <- factor(ES2005$nationality)
table(ES2005$nationality, ES2005$response)
```

No need to create a dataframe for Spain 2012, just included here for reference
```{r}
table(ES2012$nationality, ES2012$response)
```

As shown above, there were 115 atheists in 2005 and only 103 in 2012 (in the survey). We already have an inference for Spain 2012 from before but we'll do it again; We have over 10 observations for atheists in 2005 and 2012, so it's safe to use the `inference` function.

Spain in 2005:
```{r}
inference(ES2005$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
```
Textual representation of the above information, where success is considered an atheist for simplicity sake:
```{r}
## Single proportion -- success: atheist 
## Summary statistics: p_hat = 0.1003 ;  n = 1146 
## Check conditions: number of successes = 115 ; number of failures = 1031 
## Standard error = 0.0089 
## 95 % Confidence interval = ( 0.083 , 0.1177 )
```

And for 2012:
```{r}
inference(ES2012$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
```
Textual representation of the above information, where success is considered an atheist for simplicity sake:
```{r}
## Single proportion -- success: atheist 
## Summary statistics: p_hat = 0.09 ;  n = 1145 
## Check conditions: number of successes = 103 ; number of failures = 1042 
## Standard error = 0.0085 
## 95 % Confidence interval = ( 0.0734 , 0.1065 )
```

ANSWER 1C (cont.):    In conjunction with the code and relative results included above, I'll "fail to reject" the null hypothesis as the p_hat values were within the expected ranges. It's worth mentioning that the variance is not under the significance level (5% || 0.05), but based on the information given above in 1B, I have reason to believe that this change in atheism isn't representative of any trend other than normal variance... Especially given that this was over a period of 7 years.


ANSWER 2:   The confidence interval (CI) for the Spain 2012 frame is (0.0734, 0.1065), while the CI for the United States in 2012 is (0.0364, 0.0634) (Source: `Week 6 Demo Spring 2021.Rmd`). Both of these ranges were generated by a 95% CI using `inference()`. We can see clearly that the two do not overlap at all, and the higher range for Spain suggests that the population tends to lean more towards atheism. 


ANSWER 3:   Based on the formula given in the lectures, we can calculate the number of Spain residents needed to be included in a study to appropriately determine the proportion of residents who are atheists. From the lab instructions, our margin of error is `0.01%`. From the previous calculations, our p_hat is `0.09`. See the formula below, but the number of residents needed is 3,146.27; Since we can't find 27/100ths of a human, we can round up to 3,147 people to fit the criteria. 

```{r}
p_hat_2012 <- 0.09 # Using the most recent study from Spain (See Q1)
margin_error <- 0.01 # Margin of error allocated per the government
sample_size <- (1.96 ^ 2) * (p_hat_2012) * (1 - p_hat_2012) / (margin_error ^ 2) # Sample of residents needed
```

ANSWER 4:   Because the number of atheists is less than 10 (it's 8), we need to use bootstrapping to generate a larger representative sample. See below for the construction of the dataframe and bootstrapping methods.

```{r}
ZAF2012 <- subset(atheism, nationality == "South Africa" & year == "2012")
ZAF2012$nationality <- factor(ZAF2012$nationality)
table(ZAF2012$nationality, ZAF2012$response)
```

As mentioned in the lab and given by the table above, there are not 10 instances of each successes and failures (In this case, only 8 "successes" || 8 atheists). To address this, we can mimic the approach used in the `Week 6 Demo Spring 2021.Rmd` demonstration. 

We need to include the required libraries/packages:
```{r}
library(infer)
library(tidyverse)
```

Now, we'll combine knowledge from the previous Data Camp assignment and the demonstration mentioned above to create the bootstrapped samples. We'll start with creating the bootstrap sample and visualizing it with a plot.
```{r}
boot_dist <- ZAF2012 %>%
  specify(response = response, success = "atheist") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop")

ggplot(boot_dist, aes(x = stat)) +
  geom_density()
```

Moving on, now we need to find the standard error of the distribution to approximate a 95% CI.
```{r}
p_hat = 0.25
SE <- boot_dist %>%
  summarize(sd(stat)) %>%
  pull()
SE

c(p_hat - 2 * SE, p_hat + 2 * SE)
```

Using the values generated above, we can generate a plot to visualize the SE/CI.
```{r}
ggplot(boot_dist, aes(x = stat)) +
  geom_bar()
```
Finally, using the 0.025 and 0.975 percentiles to define the edges of the CI.
```{r}
quantile(boot_dist$stat, c(0.025, 0.975))
```