---
title: "Project 1"
authors: 'Team 0201 #5, Yiling Du, Bryce Hawkins, Onyekachi Igwenagu, Alec Mattu
  Delmar Randolph'
date: "03/02/2021"
output: html_document
---
**Title:** Briefing Memorandum

**To:** Dr. Random

**From:** Yiling Du, Alec Mattu, Bryce Hawkins, Onyekachi Igwenagu

**Subject:** Analysis of 110 automotive vehicle models from 2015


## Background
The intent of the analysis herein is to provide deep insights into the 110 observations compiled and provided to us by www.lock5stat.com. This compilation of 110 vehicles was provided to lock5stat by KBB.com in 2015, when these vehicles were first available for sale (new). For each of these observations (vehicles) a set of 20 variables is provided. The 110 observations consist of various vehicle models available for sale in 2015. However, this sample only includes a very small proportion of all of the vehicles available for sale in the United States in 2015, and is not representative of the entire set of vehicles on the road today. Aside from providing evaluations of various aspects about this dataset, the goal is to locate trends common between different variables for these vehicles sold in 2015. 

## Data Analyses

### Analysis 1: One Quantitative Variable
#### Analyst: Yiling Du

sign "Cars2015" to cars, and show cars
```{r}
cars <- read.csv("~/Downloads/Cars2015.csv", header=T)
cars
```
analyze the Quantitative Variable "CityMPG" (City miles per gallon), get the five number summary, mean, standard deviation, and variance.
```{r}
summary(cars$CityMPG)
sd(cars$CityMPG)
var(cars$CityMPG)

```
draw histogram and box plot for CityMPG
```{r}
hist(cars$CityMPG, breaks=5, 
     main = "Distribution of CityMPG",
     xlab= "CityMPG(City miles per gallon)",
     ylab= "Number of cars")
boxplot(cars$CityMPG,
        main = "Distribution of CityMPG",
        ylab= "CityMPG(City miles per gallon)")
```
Analysis 1 Discussion:
From the five number summary, and the mean, we can know median < mean, so the distribution is right-skewed. We also can see right-skewed from the histogram easily. We can see there is one prominent peak in the histogram, so it is unimodal. 
From the box plot, we can see there is an outlier.
Median and IQR are more robust to skewness than mean and s.d, so in this case, it is more helpful to use median and IQR.
The median/center is 20, it means 55 cars' City miles per gallon is less than 20, and 55 cars's City miles per gallon is more than 20.
IQR is good to describe the spread of the database. IQR=Q3-Q1=7, it means the range of the middle 50% of the data span is 7.


### Analysis 2: One Categorical Variable
#### Analyst: Alec Mattu

Lets begin the interpretation by loading the Cars2015.csv file into an accessible variable names cars. After we read the dataset into our environment, we just verify that the data did successfully load by typing `cars` into the console.
```{r}
cars <- read.csv("~/Downloads/Cars2015.csv", header=T)
cars
```

For this analysis of a single categorical variable, I've chosen to breakdown the vehicle type category. Although I could have picked another categorical variable, such as Size (Sm/Md/Lg) or Drive (FWD/RWD/AWD), I figured it would be neat to take a look at a variable with so many options (Small, Hatchback, Sedan, etc).

Lets begin the core portion of the analysis by listing the frequency of the categories in the dataset.
```{r}
table(cars$Type)
```

From the output generated above, we can see that there are 6 categories of vehicles: 7-Passenger, Hatchback, Sedan, Sporty, SUV, and Wagon. Of those categories, the distribution is as such: 7-Passenger (15), Hatchback (11), Sedan (46), Sporty (11), SUV (18), and Wagon (9). 

Let's take a look at a pie chart to visualize the outputs from the above table. First, we will make use of a library called Plyr, which allows us to count the frequency of each type and apply it to the `c()` function easily.

```{r}
library('plyr')
```

Then we will call the `count()` function, provided by Plyr.
```{r}
freq_table <- count(cars$Type)
freq_table
```
Now that this data is easily split up, we can apply it to the `c()` function and build a pie chart. We'll start building the pie chart by assigning a group of colors to the `label_colors` variable. This will be used for our table and graph slices.
```{r}
label_colors <- c("wheat1", "violetred4", "yellow2", "grey50", "azure1", "red")
```

Then we'll grab the frequency table data (See above) and apply the categorical values (Sedan, SUV, etc) to the variable `labels`. This will be used for our legend. After we do that, we'll grab the pie chart slice labels (The percentages per chunk) and assign that to the variable `percentageLabels`. 
```{r}
labels <- c(freq_table$x)
percentageLabels <- paste(round(freq_table$freq / 110, 1) * 100,
                          "%",
                          sep = "")
```

Just like we did for the labels, we will define the values for the slices. The only exception is that these slices are a numerical value, unlike the labels.
```{r}
slices <- c(freq_table$freq)
```

The last two bits of code we will use to build the pie chart and legend.
```{r}
pie(slices,
    labels = percentageLabels,
    main = "Vehicle Type Pie Chart",
    col = label_colors)
legend("topright",
       freq_table$x,
       fill = label_colors)
```

From this visualization above, we can easily see that the sedan class dominates the rest of the types. However, this IS a statistics class, so let's get hardcore numbers.

In order to determine the proportion of vehicles in a specific category, we need to calculate the number of vehicles listed in the dataset. A simple call to `nrows()` will give us this number.
```{r}
number_of_rows <- nrow(cars)
number_of_rows
```

This gives us 110 total rows, which we can verify by adding up the number of vehicles in each type category. Or, we can call `sum(slices)` to verify that every vehicle in the dataset has a type assigned.

```{r}
sum(slices) == number_of_rows
```

Now that we've verified that the number of rows/observations equals the number of types counted, we can calculate concrete numbers.

The number of 7-Passenger vehicles in proportion to the 110 figure we found above:
```{r}
freq_table$freq[1] / number_of_rows
```
This figure indicates that 7-pass vehicles represent ~13.6% of the vehicles apart of the dataset.

Next, we take a look at Hatchback vehicles:
```{r}
freq_table$freq[2] / number_of_rows
```
Not far off of 7-passenger vehicles, we see that hatchbacks make up 10% of the vehicles in the dataset.

Moving onto sedans, we can anticipate ~40% of the vehicles are comprised of this vehicle type; Let's find out for sure:
```{r}
freq_table$freq[3] / number_of_rows
```
No surprise here, this was present visually in the graph already. Sedans make up nearly ~42% of the 110 vehicles in the dataset.

Next to the smaller group of vehicle types, we're evaluating the "Sporty", "SUV", and "Wagon" types below.

As with the other types, take the sporty type and divide it by the sum of vehicle type observations.
```{r}
freq_table$freq[4] / number_of_rows
```
No abnormalities here, the sporty type accounts for 10% of the vehicle dataset sample.

Moving on to the SUV type,
```{r}
freq_table$freq[5] / number_of_rows
```
Although substantially under the Sedan type, the SUV comes in at just under 17% of the sample. 

Last and it just so happens to be least, the Wagon type makes up a mere 8% of the vehicle types in the dataset. Not exactly a surprise, given the rarity of finding a wagon on the road today.
```{r}
freq_table$freq[6] / number_of_rows
```

Now that we have seen a pie chart and actual representative numbers, let's take a look at the histogram to solidify our analyis.

Let's create the proper data frames to supply to the bar chart functions.
```{r}
type_df <- data.frame(x = freq_table$x, y = freq_table$freq)
type_df
```
The table above is no different than the frequency table, with the exception that it's now usable by the bar plot function with

Let's generate the bar plot to display the frequency of vehicles in a particular vehicle type category. This process is simple to facilitate because of the existing variables we just defined above.
```{r}
barplot(type_df$y, names.arg = type_df$x)
```

Of course this graph is no different in utility from the pie chart, but it allows us to see the absolute values of the vehicle types represented in a visual manner. From this, again we can see that the sedan type outnumbers all of the other types, and SUV comes in second place by a very large margin. Based on information gathered from this dataset, it's a safe assumption that passenger vehicles (sedans) make up a large proportion of vehicles on the road in The United States of America.

### Analysis 3: Two Categorical Variables
#### Analyst:[Name of Team Member]

[Put text of analysis here and add as many code chunks and text sections as you need.]

```{r}

```




### Analysis 4: One Categorical and One Quantitative Variable
#### Analyst:[Name of Team Member]

[Put text of analysis here and add as many code chunks and text sections as you need.]

```{r}

```




### Analysis 5: Two Quantitative Variable
#### Analyst:[Name of Team Member]

[Put text of analysis here and add as many code chunks and text sections as you need.]

```{r}

```





## Recommendations

[Add text section with your recommendations here. In this section, you will briefly summarize the most interesting features of your data. Are you seeing any trends, associations, or interesting features? Highlight these and make a recommendation that they be subjected to further statistical analysis. If there isn’t anything interesting, say so!]


